### 2016-03-29 Chapters/chap3/Manuscript-FINAL.tex
Solve label conflicts

One of the most important and fundamental processes in chemical  dynamics is
that of energy and charge transfer between molecular species.  The mechanism
itself is highly quantum mechanical in nature and involves strong coupling between nuclear and
electronic degrees of freedom.  The seminal model for the calculation of the transfer rate for this process
was developed by Marcus in the 1950's\cite{marcus1956theory,marcus1965theory,marcus1993electron}
\begin{equation}
k_{Marcus}=\frac{2\pi}{\hbar}|V_{ab}|^{2}\frac{1}{\sqrt{4\text{\ensuremath{\pi}}k_{B}T\lambda}}e^{-(\lambda+ \Delta G^{o})^{2}/4\text{\ensuremath{\lambda}}k_{B}T}\label{eq:marcus},
\end{equation}
which relates the Fermi golden-rule transition rate, $k_{Marcus}$
to the thermodynamic driving force $\Delta G^{o}$ and the reorganization energy $\lambda$.
One of the most striking predictions of this theory is that as the free energy difference between the final and initial state increases, the transition rate reaches its maximum,
and further increases in the energy off-set leads to slower transition rates.
This ``inverted regime''  was confirmed by Closs and Miller\cite{miller1984intramolecular,closs1988determination,closs1989connection}
nearly 30 yrs later.

@@

To calculate transition rate, we transform to adiabatic representation. It is done via Edmiston-Ruendenberg (ER) localization\cite{subotnik2010predicting} implemented in Q-Chem and the electronic Hamiltonians
on adiabatic and diabatic basis are related by the mixing angle $\theta$.
\begin{equation}
H_{dia,e}=\left(\begin{array}{cc}
\cos\theta & -\sin\theta\\
\sin\theta & \cos\theta
\end{array}\right)\left(\begin{array}{cc}
\epsilon_{1}  & 0 \\
0  & \epsilon_{2}
\end{array}\right)\left(\begin{array}{cc}
\cos\theta & \sin\theta\\
-\sin\theta & \cos\theta
\end{array}\right).\label{eq:boys}
\end{equation}
The diabatic coupling is then given by
\begin{eqnarray}
V_{12}=\frac{1}{2}\sin2\theta\left(\epsilon_{2}-\epsilon_{1}\right).
\end{eqnarray}
The full adiabatic Hamiltonian reads:
\begin{eqnarray}
H&=&U^{T}H_{dia}U\nonumber  \\
&=&\left(\begin{array}{cc}
E_{1}  & 0 \\
0 & E_{2}
\end{array}\right)+\left(\begin{array}{cc}
\sin^{2}\theta & \frac{1}{2}\sin2\theta\\
\frac{1}{2}\sin2\theta & \cos^{2}\theta
\end{array}\right) {\mathbf g}_{22}.{\mathbf q} \nonumber \\
&+& H_{osc}.\label{eq:locaiHam}
\end{eqnarray}
where the electronic part is in the adiabatic basis with eigenenergies of $E_1$ and $E_2$ at certain reference geometry. The nuclear normal modes are described by $H_{osc}$ whereas $\theta$ is the mixing angle between two states.
%to describe the coupling between two electronic states, $|1\rangle$ and $|2\rangle$  by a set of nuclear vibrational motionsdescribed by $H_{osc}$.  The angle $\theta$ gives the mixing angle between diabatic states and $E_1$ and $E_{2}$ are electronic energies computed at the same reference configuration.

To parametrize our Hamiltonian, we assume the diabatic
are a good approximation to the actual adiabatic potentials.
This allows us to
use the gradients of the adiabatic potentials to approximate the diabatic potentials.
This approximation is  valid when the adiabatic (and diabatic) energy minima
are far enough away from the crossing points and the mixing angles between the diabatic  and adiabatic
states is small.
Thus, all parameters in Eq.~\ref{eq:locaiHam}  can be  obtained from
standard  quantum chemical computations.
As in our previous work, we use the Q-Chem 4.0 package  to obtain the vertical energies
 from single point CI(S) calculations with 6-31G** basis set at a given
reference geometry.   We then project the energy gradients onto the vibrational normal
coordinates

Our group has shown that the dynamics of Hamiltonian in the form of Eq.~\ref{eq:locaiHam} can be expressed in a time-convolutionless mater equation \cite{pereverzev2006time}. The derivation is lengthy so we present it briefly.

\subsection{Transition Rate and Autocorrelation}
We start from a more general Hamiltonian:
\begin{eqnarray}
H=\left(\begin{array}{cc}
\epsilon_{1} & 0\\
0 & \epsilon_{2}
\end{array}\right)+\left(\begin{array}{cc}
{\mathbf g}_{11}&{\mathbf g}_{12} \\
{\mathbf g}_{21} &{\mathbf g}_{22}
\end{array}\right)\cdot{\mathbf q} +\frac{{\mathbf p}^{2}}{2}+\frac{1}{2}\mathbf{q}^{T}\cdot\mathbf\Omega\cdot\mathbf{q}
\nonumber \\
\label{ham1}
\end{eqnarray}
then perform a polaron transform
using the unitary transformation~\cite{grover1970exciton,rice1994excitons,pereverzev2006time}.
\begin{eqnarray}
U&=&e^{-\sum_{ni}\!\!\frac{g_{nni}}{\hbar\omega_i}|n\rangle \langle
n|(a^{\dagger}_i-a_i)}
 \nonumber \\
&=&
\sum_{n}|n\rangle \langle n|e^{-\sum_{i}\!\!\frac{g_{nni}}{\hbar\omega_i}(a^{\dagger}_i-a_i)}
\label{unitary}
\end{eqnarray}
under which the transformed Hamiltonian is written in terms of the
diagonal elements
\begin{eqnarray} \tilde H_0=U^{-1}H_0U
=\sum_n\tilde\epsilon_n |n\rangle \langle
n|+\sum_i\omega_ia^{\dagger}_ia_i,
 \end{eqnarray}
with  the renormalized electronic energies,
\begin{eqnarray}
\tilde\epsilon_n=\epsilon_n-\sum_{i}\frac{g_{nni}^2}{\hbar\omega_i},
\end{eqnarray}
and off-diagonal terms,
\begin{eqnarray} \hat V_{nm}=\sum_{i}g_{nmi}\left(a^{\dagger}_i+
a_i-\frac{2g_{nni}}{\hbar\omega_i}\right)e^{\sum_{j}\frac{(g_{nnj}-g_{mmj})}{\hbar\omega_j}(a^{\dagger}_j-a_j)}.
\label{opm}
\end{eqnarray}
% In the transformed (or dressed) picture the electronic transition from state
% $|n\rangle$ to $|m\rangle$ is accompanied by the excitations of all the
% normal phonon modes.  Transforming to the interaction representation
% and performing a trace over the phonons gives the spectral density in
% terms of the autocorrelation of the electron/phonon coupling
% operators.
% \begin{eqnarray}
% S_{nm}(\tilde\omega) = \int_{-\infty}^{\infty} dt e^{-i\tilde \omega t}C_{nm}(t),\label{spec-dens}
% \end{eqnarray}
% where
% $$
%C_{nm}(t) = \int_{-\infty}^{\infty} dt e^{-i\tilde \omega t}\langle \hat V_{nm}(t) \hat V_{mn}(0)\rangle.\label{cor-fun}
After lengthy derivation we get the autocorrelation function in the Heisenberg representation
$$
C_{nm}(t) =\langle \hat V_{nm}(t) \hat V_{mn}(0)\rangle,\label{cor-fun}
$$
where $\langle \cdots \rangle$ denotes a thermal average over the
vibrational degrees of freedom.
% The derivation and explicit form for the kernel in Eq.~\ref{spec-dens}  is quite lengthy and is given in
% Ref.~\citenum{pereverzev2006time}.
Using this approach , the golden-rule rates are given by
\begin{eqnarray}
k_{nm}=\lim_{\tau \to \infty}2{\rm Re}\int_{0}^{\tau}dt\left\langle \hat V_{nm}(0)\hat V_{mn}\left(t\right)\right\rangle e^{-i\tilde\omega_{nm}t}.
\label{gr-expression}
\end{eqnarray}
In a practical sense, we take $\tau$ to be some finite time at which the
autocorrelation function $C(t)$
has relaxed to zero.

@@

\subsection{Constructing Primary Lanczos Modes}
The linear coupling parameter $ {\mathbf g}_{22}$ in Eq.~\ref{eq:locaiHam} defines
a force
along a vector connecting
the initial and final equilibrium geometries of the molecule.
This vector, along with the diabatic mixing angle can be obtained from quantum chemistry
using the ER localization
scheme to approximate the donor and acceptor states.
By analyzing this force we can gain the insight into the dynamics of the transition as well as
open an avenue for developing improved approximations for transition rates.
In our previous work \cite{yang2014intramolecular},  we presented a Lanczos-base ranking algorithm that project
out a series of nuclear displacements that are most important for
the transition.   We refer to the highest-ranked mode identified by the algorithm
as the ``primary Lanczos mode'' (PLM).
The process of finding the PLMs is initiated by defining the vector ${\mathbf v}_{1} = {\mathbf g}_{22}$.
At each step indexed by $k$,  we define a projection operator
\begin{eqnarray}
{\mathbf P}_{k} = {\mathbf v}_{k}\otimes  {\mathbf v}_{k}
\end{eqnarray}
and its complement ${\mathbf Q}_{k} = {\mathbf I}  - {\mathbf P}_{k}$.
We also construct
\begin{eqnarray}
\mathbf p = \sum_{k} \mathbf P_{k}
\end{eqnarray}
as  the total projection operator for all ${k} \le N$ modes.
We then project the hessian matrix $\mathbf\Omega$ into each subspace {\em viz.}
\begin{eqnarray}
\mathbf\Omega_{p} = \mathbf P_{k}\cdot \mathbf\Omega \cdot \mathbf P_{k} \,\, \& \,\, \mathbf\Omega_{q} = \mathbf Q_{k}\cdot \mathbf\Omega \cdot  \mathbf Q_{k}
\end{eqnarray}
and diagonalize each to obtain eigenvalues and eigenvectors $\{\alpha_{p}, {\mathbf M}_{p}\}$ and $\{\alpha_{q}, {\mathbf M}_{q}\}$
respectively.  $\mathbf\Omega_{p} $ and $\mathbf\Omega_{q}$ are $N\times N$ matrices.
The first set of these  will have a single
non-trivial eigenvalue and the second set
will have $N-k$ non-trivial eigenvalues. We then collect the non-trivial eigenvectors associated with each
to form the orthogonal transformation matrix
\begin{eqnarray}
{\mathbf M}_{k} = \{{\mathbf M}_{p},{\mathbf M}_{q}\}.
\end{eqnarray}
and again transform the full hessian $\mathbf\Omega$ into this new vector space to form the $N\times N$ matrix $\mathbf\Omega'$.
 At each step in the iteration, the transformed hessian, $\mathbf\Omega'$ is in the form of a
$k\times k$ tri-diagonal submatrix in the upper-left part of the matrix and
a diagonal submatrix in the lower-right.

>>>>

One of the most important and fundamental processes in chemical  dynamics is
that of energy and charge transfer between molecular species.  The mechanism
itself is highly quantum mechanical in nature and involves strong coupling between nuclear and
electronic degrees of freedom.  The seminal model for the calculation of the transfer rate for this process
was developed by Marcus in the 1950's\cite{marcus1956theory,marcus1965theory,marcus1993electron}
\begin{equation}
k_{Marcus}=\frac{2\pi}{\hbar}|V_{ab}|^{2}\frac{1}{\sqrt{4\text{\ensuremath{\pi}}k_{B}T\lambda}}e^{-(\lambda+ \Delta G^{o})^{2}/4\text{\ensuremath{\lambda}}k_{B}T},
\end{equation}
which relates the Fermi golden-rule transition rate, $k_{Marcus}$
to the thermodynamic driving force $\Delta G^{o}$ and the reorganization energy $\lambda$.
One of the most striking predictions of this theory is that as the free energy difference between the final and initial state increases, the transition rate reaches its maximum,
and further increases in the energy off-set leads to slower transition rates.
This ``inverted regime''  was confirmed by Closs and Miller\cite{miller1984intramolecular,closs1988determination,closs1989connection}
nearly 30 yrs later.

@@

To calculate transition rate, we transform to adiabatic representation. It is done via Edmiston-Ruendenberg (ER) localization\cite{subotnik2010predicting} implemented in Q-Chem and the electronic Hamiltonians
on adiabatic and diabatic basis are related by the mixing angle $\theta$.
\begin{equation}
H_{dia,e}=\left(\begin{array}{cc}
\cos\theta & -\sin\theta\\
\sin\theta & \cos\theta
\end{array}\right)\left(\begin{array}{cc}
\epsilon_{1}  & 0 \\
0  & \epsilon_{2}
\end{array}\right)\left(\begin{array}{cc}
\cos\theta & \sin\theta\\
-\sin\theta & \cos\theta
\end{array}\right).
\end{equation}
The diabatic coupling is then given by
\begin{eqnarray}
V_{12}=\frac{1}{2}\sin2\theta\left(\epsilon_{2}-\epsilon_{1}\right).
\end{eqnarray}
The full adiabatic Hamiltonian reads:
\begin{eqnarray}
H&=&U^{T}H_{dia}U\nonumber  \\
&=&\left(\begin{array}{cc}
E_{1}  & 0 \\
0 & E_{2}
\end{array}\right)+\left(\begin{array}{cc}
\sin^{2}\theta & \frac{1}{2}\sin2\theta\\
\frac{1}{2}\sin2\theta & \cos^{2}\theta
\end{array}\right) {\mathbf g}_{22}.{\mathbf q} \nonumber \\
&+& H_{osc}.\label{eq:locaiHam2}
\end{eqnarray}
where the electronic part is in the adiabatic basis with eigenenergies of $E_1$ and $E_2$ at certain reference geometry. The nuclear normal modes are described by $H_{osc}$ whereas $\theta$ is the mixing angle between two states.
%to describe the coupling between two electronic states, $|1\rangle$ and $|2\rangle$  by a set of nuclear vibrational motionsdescribed by $H_{osc}$.  The angle $\theta$ gives the mixing angle between diabatic states and $E_1$ and $E_{2}$ are electronic energies computed at the same reference configuration.

To parametrize our Hamiltonian, we assume the diabatic
are a good approximation to the actual adiabatic potentials.
This allows us to
use the gradients of the adiabatic potentials to approximate the diabatic potentials.
This approximation is  valid when the adiabatic (and diabatic) energy minima
are far enough away from the crossing points and the mixing angles between the diabatic  and adiabatic
states is small.
Thus, all parameters in Eq.~\ref{eq:locaiHam2}  can be  obtained from
standard  quantum chemical computations.
As in our previous work, we use the Q-Chem 4.0 package  to obtain the vertical energies
 from single point CI(S) calculations with 6-31G** basis set at a given
reference geometry.   We then project the energy gradients onto the vibrational normal
coordinates

Our group has shown that the dynamics of Hamiltonian in the form of Eq.~\ref{eq:locaiHam2} can be expressed in a time-convolutionless mater equation \cite{pereverzev2006time}. The derivation is lengthy so we present it briefly.

\subsection{Transition Rate and Autocorrelation}
We start from a more general Hamiltonian:
\begin{eqnarray}
H=\left(\begin{array}{cc}
\epsilon_{1} & 0\\
0 & \epsilon_{2}
\end{array}\right)+\left(\begin{array}{cc}
{\mathbf g}_{11}&{\mathbf g}_{12} \\
{\mathbf g}_{21} &{\mathbf g}_{22}
\end{array}\right)\cdot{\mathbf q} +\frac{{\mathbf p}^{2}}{2}+\frac{1}{2}\mathbf{q}^{T}\cdot\mathbf\Omega\cdot\mathbf{q}
\nonumber \\
% \label{ham1}
\end{eqnarray}
then perform a polaron transform
using the unitary transformation~\cite{grover1970exciton,rice1994excitons,pereverzev2006time}.
\begin{eqnarray}
U&=&e^{-\sum_{ni}\!\!\frac{g_{nni}}{\hbar\omega_i}|n\rangle \langle
n|(a^{\dagger}_i-a_i)}
 \nonumber \\
&=&
\sum_{n}|n\rangle \langle n|e^{-\sum_{i}\!\!\frac{g_{nni}}{\hbar\omega_i}(a^{\dagger}_i-a_i)}
% \label{unitary}
\end{eqnarray}
under which the transformed Hamiltonian is written in terms of the
diagonal elements
\begin{eqnarray} \tilde H_0=U^{-1}H_0U
=\sum_n\tilde\epsilon_n |n\rangle \langle
n|+\sum_i\omega_ia^{\dagger}_ia_i,
 \end{eqnarray}
with  the renormalized electronic energies,
\begin{eqnarray}
\tilde\epsilon_n=\epsilon_n-\sum_{i}\frac{g_{nni}^2}{\hbar\omega_i},
\end{eqnarray}
and off-diagonal terms,
\begin{eqnarray} \hat V_{nm}=\sum_{i}g_{nmi}\left(a^{\dagger}_i+
a_i-\frac{2g_{nni}}{\hbar\omega_i}\right)e^{\sum_{j}\frac{(g_{nnj}-g_{mmj})}{\hbar\omega_j}(a^{\dagger}_j-a_j)}.
% \label{opm}
\end{eqnarray}
% In the transformed (or dressed) picture the electronic transition from state
% $|n\rangle$ to $|m\rangle$ is accompanied by the excitations of all the
% normal phonon modes.  Transforming to the interaction representation
% and performing a trace over the phonons gives the spectral density in
% terms of the autocorrelation of the electron/phonon coupling
% operators.
% \begin{eqnarray}
% S_{nm}(\tilde\omega) = \int_{-\infty}^{\infty} dt e^{-i\tilde \omega t}C_{nm}(t),\label{spec-dens}
% \end{eqnarray}
% where
% $$
%C_{nm}(t) = \int_{-\infty}^{\infty} dt e^{-i\tilde \omega t}\langle \hat V_{nm}(t) \hat V_{mn}(0)\rangle.\label{cor-fun}
After lengthy derivation we get the autocorrelation function in the Heisenberg representation
$$
C_{nm}(t) =\langle \hat V_{nm}(t) \hat V_{mn}(0)\rangle,\label{cor-fun}
$$
where $\langle \cdots \rangle$ denotes a thermal average over the
vibrational degrees of freedom.
% The derivation and explicit form for the kernel in Eq.~\ref{spec-dens}  is quite lengthy and is given in
% Ref.~\citenum{pereverzev2006time}.
Using this approach , the golden-rule rates are given by
\begin{eqnarray}
k_{nm}=\lim_{\tau \to \infty}2{\rm Re}\int_{0}^{\tau}dt\left\langle \hat V_{nm}(0)\hat V_{mn}\left(t\right)\right\rangle e^{-i\tilde\omega_{nm}t}.
% \label{gr-expression}
\end{eqnarray}
In a practical sense, we take $\tau$ to be some finite time at which the
autocorrelation function $C(t)$
has relaxed to zero.

@@

\subsection{Constructing Primary Lanczos Modes}
The linear coupling parameter $ {\mathbf g}_{22}$ in Eq.~\ref{eq:locaiHam2} defines
a force
along a vector connecting
the initial and final equilibrium geometries of the molecule.
This vector, along with the diabatic mixing angle can be obtained from quantum chemistry
using the ER localization
scheme to approximate the donor and acceptor states.
By analyzing this force we can gain the insight into the dynamics of the transition as well as
open an avenue for developing improved approximations for transition rates.
In our previous work \cite{yang2014intramolecular},  we presented a Lanczos-base ranking algorithm that project
out a series of nuclear displacements that are most important for
the transition.   We refer to the highest-ranked mode identified by the algorithm
as the ``primary Lanczos mode'' (PLM).
The process of finding the PLMs is initiated by defining the vector ${\mathbf v}_{1} = {\mathbf g}_{22}$.
At each step indexed by $k$,  we define a projection operator
\begin{eqnarray}
{\mathbf P}_{k} = {\mathbf v}_{k}\otimes  {\mathbf v}_{k}
\end{eqnarray}
and its complement ${\mathbf Q}_{k} = {\mathbf I}  - {\mathbf P}_{k}$.
We also construct
\begin{eqnarray}
\mathbf p = \sum_{k} \mathbf P_{k}
\end{eqnarray}
as  the total projection operator for all ${k} \le N$ modes.
We then project the hessian matrix $\mathbf\Omega$ into each subspace {\em viz.}
\begin{eqnarray}
\mathbf\Omega_{p} = \mathbf P_{k}\cdot \mathbf\Omega \cdot \mathbf P_{k} \,\, \& \,\, \mathbf\Omega_{q} = \mathbf Q_{k}\cdot \mathbf\Omega \cdot  \mathbf Q_{k}
\end{eqnarray}
and diagonalize each to obtain eigenvalues and eigenvectors $\{\alpha_{p}, {\mathbf M}_{p}\}$ and $\{\alpha_{q}, {\mathbf M}_{q}\}$
respectively.  $\mathbf\Omega_{p} $ and $\mathbf\Omega_{q}$ are $N\times N$ matrices.
The first set of these  will have a single
non-trivial eigenvalue and the second set
will have $N-k$ non-trivial eigenvalues. We then collect the non-trivial eigenvectors associated with each
to form the orthogonal transformation matrix
\begin{eqnarray}
{\mathbf M}_{k} = \{{\mathbf M}_{p},{\mathbf M}_{q}\}.
\end{eqnarray}
and again transform the full hessian $\mathbf\Omega$ into this new vector space to form the $N\times N$ matrix $\mathbf\Omega'$.
 At each step in the iteration, the transformed hessian, $\mathbf\Omega'$ is in the form of a
$k\times k$ tri-diagonal submatrix in the upper-left part of the matrix and
a diagonal submatrix in the lower-right.
